Zohmg---a large scale data warehouse for aggregated time-series-based data. [DRAFT]

### 2009-05-05 doc/report/msc-report.tex
doc/report/msc-report.tex: + merge of Hadoop, HBase and Dumbo sections. + additions to Apache/Thrift, Serializing, Java and Python.

\author{Per Andersson, Fredrik M{\"o}llerstrand}
\title{Zohmg--a large scale data warehouse for aggregated time-series-based
data. [DRAFT]}

@@

All three requirements are depend on one another, what data and how it is
stored affects what can be visualised. Also, what data is imported--and
how it is analysed--affects how it can be stored and hence visualised.

@@

Assuming that time-series analysis is the most important, we have imposed
the restriction on the system that is built so that it only handles
time-series. Because of this restriction the data store obviously becomes
less general purpose. Although since most log data is time-series based
this restriction does not cripple the system much--if at all.

@@

The data model is a simple one: Rows of data are stored in named tables.
Each row consists of a sortable row key and an arbitrary number of columns.
The columns are of the form "column-family:qualifier", where the column
family is one of a number of fixed column families defined by the table,
and the qualifier is an arbitrary string specified by the application. In
effect, column families are pre-announced while qualifiers are not. The
contents of each column family are stored together, so the user will want
to store items that have similar characteristics in the same column family.
Each data point is called a cell. A cell is an uninterpreted array of
bytes-- there are no data types, and it is up to the application to
interpret the data correctly. The table can store any number of timestamped
versions of each cell, allowing versioning of data.

@@

In actual practice, the main difference between the two models is that
querying is highly restricted in a BigTable-like data store--there are no
joins, for example--whereas the relational database was born to be queried.
On the other hand, the restrictions placed upon the BigTable models are
what allows it to scale to such a large dataset. Indeed, BigTable was born
to scale.

@@

A region is a partition of the data stored on a node in the computer
cluster. A region server (referred to as tablet server in \cite{mapreduce})
holds information on which regions are located on which nodes. More
precisely the region server knows what range of rowkeys are in each region
on each node. This behaviour makes it fast to access data while keeping
reliability and scalability features.

@@

Hadoop is an open-source implementation of Google's MapReduce-framework described in [mapreduce]. Hadoop is written in Java and is part of the Apache Software Foundation.  It is a fairly mature piece of software and is used in production at numerous companies.

@@

All Zohmg programs are executed on Hadoop.

@@

HBase is an open-source implementation of Google's BigTable data store described in [bigtable]. It is a sub-project of Hadoop, written in Java and a part of the Apache Software Foundation. HBase is still in its infancy and few organizations use it for mission-critical applications.

@@

Sparseness is a big thing. Null values are free of charge. *Describe how we leverage sparsity*.

@@

@@

Dumbo is a framework for writing Hadoop programs in languages other than Java, in a so-called Streaming Mode. Dumbo is written in Python. It is used extensively at last.fm for writing short prototypes.

@@

@@

@@

@@

@@

@@

\section{JSON}

@@

OHM is built upon existing tools and software. It uses Apache Hadoop and
HBase, free implementations of Google's MapReduce and BigTable
respectively, both implemented in Java.

Importing data to OHM is as simple as the user defines a mapper--the first
half of a MapReduce program--which is run over a file that resides on
HDFS\footnote{Hadoop Distributed Filesystem}.

Data is stored on Apache HBase

Exporting data

JSON\footnote{JavaScript Object Notation} is transported over a RESTful
interface or data is transported with Thrift.

@@

The user's datasets are defined using YAML--a human friendly data
serialization standard. The dataset files reside in the config directory.

@@

We have built a system--henceforth called OHM--which has the goal of being
a large scale aggregating data warehouse for time-series-based log-like
data sources.

@@

Built with Python Paste--a middleware framework--the data server dispatches
to the requested application depending on the URI it was fed.

@@

Update tables on configuration--dataset files--changes.

>>>>

\author{Per Andersson, Fredrik M{\"o}llerstrand}
\title{Zohmg---a large scale data warehouse for aggregated time-series-based
data. [DRAFT]}

@@

All three requirements are depend on one another, what data and how it is
stored affects what can be visualised. Also, what data is imported---and
how it is analysed---affects how it can be stored and hence visualised.

@@

Assuming that time-series analysis is the most important, we have imposed
the restriction on the system that is built so that it only handles
time-series. Because of this restriction the data store obviously becomes
less general purpose. Although since most log data is time-series based
this restriction does not cripple the system much---if at all.

@@

The data model is a simple one: Rows of data are stored in named tables.
Each row consists of a sortable row key and an arbitrary number of columns.
The columns are of the form "column-family:qualifier", where the column
family is one of a number of fixed column families defined by the table,
and the qualifier is an arbitrary string specified by the application. In
effect, column families are pre-announced while qualifiers are not. The
contents of each column family are stored together, so the user will want
to store items that have similar characteristics in the same column family.
Each data point is called a cell. A cell is an uninterpreted array of
bytes---there are no data types, and it is up to the application to
interpret the data correctly. The table can store any number of timestamped
versions of each cell, allowing versioning of data.

@@

In actual practice, the main difference between the two models is that
querying is highly restricted in a BigTable-like data store---there are no
joins, for example---whereas the relational database was born to be queried.
On the other hand, the restrictions placed upon the BigTable models are
what allows it to scale to such a large dataset. Indeed, BigTable was born
to scale.

@@

A region is a partition of the data stored on a node in the computer
cluster. A region server (referred to as tablet server in \cite{bigtable})
holds information on which regions are located on which nodes. More
precisely the region server knows what range of rowkeys are in each region
on each node. This behaviour makes it fast to access data while keeping
reliability and scalability features.

@@

Hadoop is an free software implementation of Google's MapReduce-framework described in \cite{mapreduce}. Hadoop is written in Java and is part of the Apache Software Foundation.  It is a fairly mature piece of software and is used in production at numerous companies.

@@

All Zohmg import programs are executed on Hadoop.

@@

HBase is an free software implementation of Google's BigTable data store described in \cite{bigtable}. It is a sub-project of Hadoop, written in Java and a part of the Apache Software Foundation. HBase is still in its infancy and few organizations use it for mission-critical applications.

@@

Sparseness is a big thing. Null values are free of charge. TODO: *Describe how we leverage sparsity*.

@@

A remote procedure call framework for building scalable cross-language
services. Thrift combines a software stack with a code generation engine to
build services that work seamlessly between a wide range of programming and
scripting languages. \cite{thrift}

Thrift was originally developed at Facebook and in April 2007 it was
released as free software. It entered the Apache Incubator in May 2008.

\subsubsection{Python}

HBase has a Thrift server which serves as the interface to languages other
than Java--for instance Python.

@@

Dumbo is a framework for writing Hadoop programs in languages other than Java, in a so-called Streaming Mode. Dumbo is written in Python. It is used extensively at Last.fm for writing short prototypes.

@@

\section{Serializing}

Serializing data is the task of converting an object into a sequence of
bits so that it can be stored on a storage medium or transmitted over a
network link. Rereading the sequence of bits should restore the original
state of the object it was in before the serialization.

\subsection{JSON}

JSON\footnote{JavaScript Object Notation} is a lightweight text-based
human-readable data interchange format. It is used for representing simple
data structures and associative arrays (called objects). JSON is often used
for serializing structured data and transmitting it over network
connection.

YAML\footnote{YAML Ain't a Markup Language} is a human-readable data
serialization format available for a wide range of programming and
scripting languages.

The outline and markup of YAML documents makes it well suited for data
which humans are likely to view or edit, such as configuration files.

@@

Both Hadoop and HBase are written in Java---a popular high-level
object-oriented programming language.

Most Hadoop users write their MapReduce programs in Java, since this is the
native language of Hadoop.

@@

Python is a popular general-purpose high-level multi-paradigm---most
notably: object-oriented, imperative and functional---scripting language.

@@

Middleware. WSGI. HTTP server.

@@

PyYAML is a complete YAML 1.1 compliant parser and emitter for Python.

@@

REST\footnote{Representational State Transfer}

@@

As stated in chapter 3, Zohmg is built upon existing tools and software. It
uses Apache Hadoop and HBase, free implementations of Google's MapReduce
and BigTable respectively, both implemented in Java.

Importing data to Zohmg is as simple as the user defines a mapper---the
first half of a MapReduce program---which is run over a file that resides
on HDFS\footnote{Hadoop Distributed Filesystem}.

Data is stored in Apache HBase.

Exporting data from Zohmg is done via a RESTful interface which dumps JSON.
It is possible to transform the stored data with user defined transformers.

@@

The user's datasets are defined using YAML. The dataset files reside in the
config directory.

@@

We have built a system---henceforth called OHM---which has the goal of being
a large scale aggregating data warehouse for time-series-based log-like
data sources.

@@

Built with Python Paste---a middleware framework---the data server dispatches
to the requested application depending on the URI it was fed.

@@

Update tables on configuration---dataset files---changes.
